---
title: "Introduction to R shiny package"
author: Hakan Turgay & Bertan Taylan
date: October 19, 2019
runtime: shiny
output:
  ioslides_presentation: 
    css: styles.css
    highlight: pygments
    incremental: yes
    logo: Rlogo.png
    transition: faster
    widescreen: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is Shiny package?

- Shiny is an R package that makes it easy to build <span style="color:red; font-weight:bold">interactive web applications</span> (apps) straight from R. This lesson will get you started building Shiny apps right away.
- Build useful web applications <span style="color:purple; font-weight:bold">with only a few lines of code</span>—no JavaScript required.
- Shiny applications are <span style="color:green; font-weight:bold">automatically “live”</span> in the same way that spreadsheets are live. Outputs change instantly as users modify inputs, without requiring a reload of the browser.
- Shiny user interfaces can be built entirely using R, or can be written directly in HTML, CSS, and JavaScript for more <span style="color:yellow; font-weight:bold">flexibility.</span>
- Pre-built output widgets for displaying <span style="color:aqua; font-weight:bold">plots, tables, and printed output of R objects.</span>

## Install Shiny Package 

If you still haven’t installed the Shiny package, open an R session, connect to the internet, and run

```r
install.packages("shiny")
```

## Say Hello Shiny

```r
library(shiny)
runExample("01_hello")
```

```{r}
library(shiny)
library(shiny)

# Define UI for app that draws a histogram ----
ui <- fluidPage(

  # App title ----
  titlePanel("Hello Shiny!"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "distPlot")

    )
  )
)




# Define server logic required to draw a histogram ----
server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Waiting time to next eruption (in mins)",
         main = "Histogram of waiting times")

    })

}

shinyApp(ui = ui, server = server)

```

## More Example

```r 
runExample("01_hello")      # a histogram
runExample("02_text")       # tables and data frames
runExample("03_reactivity") # a reactive expression
runExample("04_mpg")        # global variables
runExample("05_sliders")    # slider bars
runExample("06_tabsets")    # tabbed panels
runExample("07_widgets")    # help text and submit buttons
runExample("08_html")       # Shiny app built from HTML
runExample("09_upload")     # file upload wizard
runExample("10_download")   # file download wizard
runExample("11_timer")      # an automated timer
```

## Structure of a Shiny App

* Shiny apps are contained in a single script called app.R. The script app.R lives in a directory (for example, newdir/) and the app can be run with runApp("newdir").

* a user interface object (<span style="color:red; font-weight:bold">ui.r</span>)

* a server function       (<span style="color:red; font-weight:bold">server.r</span>)

* a call to the shinyApp function

## 